\singlespacing \vspace{-1.6cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2022-04-15 Manuscript/Lanuza_et_al_2021_Ecol_Letters.Rmd
edits
Working on 1st paragraph

Despite the astonishing diversity of floral structures among flowering plants [@barrett2002; @schiestl2013] and their importance in shaping plant-pollinator interactions [@fenster2004; @dellinger2020], a unified framework that describes the major plant reproductive trade-offs is currently lacking [@roddy2021]. In addition, macroecological studies that investigate plant reproductive traits are scarce [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017] and consequently, there is poor understanding of how reproductive traits drive interactions with pollinators at large scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations [@dehling2016]. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016]. However, plant reproductive traits have been overlooked beyond highly specialised pollination systems [@dellinger2020] and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations [@tur2013; @devaux2014].

@@

Here, we aim to explore the potential trade-offs among reproductive traits and how these influence plant-pollinator interactions. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species’ position in trait-space influence interactions with floral visitors. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species’ roles within networks using a set of complementary interaction network metrics (i.e., visitation rate, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks. These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies [@olesen2007; @fortuna2010; @carvalheiro2014] or available in online archives [e.g., The Web of Life, @fortuna2010; and Mangal, @poisot2016]. In total, our network dataset (Table S1) constituted 60 weighted (visitation rate) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative and quantitative form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with visitation rate that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and visitation rate does not consider each pollinator species efficiency [@ballantyne2015], visitation rate can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version _4.0.3_.

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and visitation rate of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that _Apis mellifera_ was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that _A. mellifera_ was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of _A. mellifera_ on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and visitation rate excluding _A. mellifera_. We found that _A. mellifera_, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package _brms_ [@burkner2017]. We modelled presence-absence of observed interactions and visitation rate as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., visitation rate ~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and visitation rate a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with weakly informative priors from the brms function, 3,000 iterations and with previous 1,000 warm up iterations. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the _bayesplot_ package [@gabry2019].

Second, to better understand complex trait relationships, we used regression trees. Regression trees are recursive algorithms which can detect complex relationships among predictors and allow identification of the relevance of specific trait combinations on explaining species roles within the network of interaction. We focused exclusively on quantitative traits because almost all categorical traits were statistically associated with the first two axes of trait variation (Table S2). We conducted this analysis using the _rpart_ package [@therneau2015] with method _‘anova’_ with a minimum of 50 observations per terminal node and we used the _rpart.plot_ package [@milborrow2015] to plot the regression trees. We considered the species level indices as response variables (visitation rate, normalized degree and specialization) and we performed one regression tree per metric using the different quantitative traits as predictors. We calculated two regression trees per plant species-level metric, one for the full set of species and another for the subset of species for which we had pollen and nectar traits. We focused on regression trees that included floral rewards because they consistently showed pollen and nectar traits as being the best for explaining the different species-level metrics (Fig. S4).

@@

\vspace{5mm}

The main axes of trait variation explained partly presence-absence of interaction partners (conditional $R^{2}$ = 0.26; marginal $R^{2}$ = 0.20) but little of the overall visitation rates (conditional $R^{2}$ = 0.31; marginal $R^{2}$ = 0.06). However, we found relevant differences across the different floral visitor guilds on both presence-absence of interactions and visitation rates (Fig. 3). We found that plants with high flower number and small flowers had higher interaction partners of Coleoptera, non-bee-Hymenoptera and all Diptera guilds but plants with low flower number but large flowers had higher interaction partners of bees and Lepidoptera guilds (flower number - flower size trade-off, PC1; Fig. 3a). However, plant species with high flower number but small flowers had higher visitation rates of bees and syrphids guilds (PC1; Fig. 3d). Remarkably, all plant species with higher pollinator dependence had higher number of interacting partners and visitation rates for all floral visitor guilds (PC2; Fig. 3b and Fig. 3e). Finally, plant species with short styles and low selfing had higher interaction partners of all guilds but bees that interacted clearly more with plant species with long styles and high selfing (style length trade-off; Fig. 3c). However, for visitation rate, we found that plants with long styles and high selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera guilds (Fig. 3f).

The additional model for both presence-absence of interaction (marginal $R^{2}$ = 0.29; conditional $R^{2}$ = 0.19) and visitation rate (marginal $R^{2}$ = 0.30; conditional $R^{2}$ = 0.03) for the most represented families of bees showed that the family Apidae was the main driver of the observed patterns. The contrasting differences between presence-absence of interaction and visitation rate for bees on PC1 (Fig. 3a and Fig. 3d) were driven by the family Andrenidae that had higher number of interacting partners but lower visitation rates on plant species with low number of large flowers (Fig. S8).

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.cap= "**Figure 3** Fitted posterior estimates of the presence-absence of interaction (a, b and c) and visitation rate (d, e and f) of the different floral visitor guilds in relation to the main axes of trait variation (PC1, PC2 and PC3). PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off. For visualization purposes, due to large differences between the visitation rates of bees and the rest of guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

The variance of the different plant species-level network metrics was poorly explained by the three main axes of trait variation (Fig. S9; visitation rate ~ PCs, conditional $R^{2}$ = 0.11, marginal $R^{2}$ = 0.02; normalized degree ~ PCs, conditional $R^{2}$ = 0.24, marginal $R^{2}$ = 0.02; and, specialization ~ PCs, conditional $R^{2}$ = 0.37, marginal$R^{2}$ = 0.03). Overall, the most notable trends were found on PC1 and PC3 for visitation rate and specialization. On the flower number - flower size trade-off (PC1), visitation rate was higher for plant species with more flowers but was lower for plant species with larger flowers (Fig. S9a). On PC1, specialization showed the opposite trend (Fig. S9g). On the style length trade-off (PC3), visitation rate was lower for plants with shorter styles and lower autonomous selfing and higher for species with longer styles and higher autonomous selfing (Fig. S9c). Again, specialization showed the opposite trend to visitation rate (Fig. S9i).

When we further investigated the combination of traits that drive plant network roles, we found that the regression tree for visitation rate was best explained by plant height, nectar concentration and style length (Fig. 4a). Specifically, species taller than 3.9m had the highest visitation rate, while species that were shorter than 3.9m and had a nectar concentration lower than 16% had the lowest visitation rate. Normalized degree was best explained by nectar concentration, pollen grains per flower, plant height, flower width and autonomous selfing (Fig. 4b). Species with a nectar concentration over 49% had the highest levels of normalized degree, whereas species with nectar concentration lower than 49%, more than 21,000 pollen grains per flower and height less than 0.78m had the lowest normalized degree. Finally, specialization was best explained by plant height, ovule number, pollen grains per flower and autonomous selfing (Fig. 4c). Overall, plant species with the highest specialization were shorter than 1.3m, had more than 14,000 pollen grains per flower and autonomously self-pollinated less than 11% of their fruits. In contrast, species taller or equal than 5.1m and with lower than 14 ovules per flower had the lowest specialization values.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.cap= "**Figure 4** Contribution of traits in plant's network roles. Regression tree analysis of visitation rate (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

@@

rpart.plot(tree1, box.palette="GnOr", main="(a) Visitation rate", cex.main=1.75,tweak = 1.25,type = 2) 
rpart.plot(tree2, box.palette="GnOr", main="(b) Normalized degree", cex.main=1.75,tweak = 1.25,type = 2)
rpart.plot(tree3, box.palette="GnOr", main="(c) Specialization", cex.main=1.75,tweak = 1.25,type = 2)

@@

This study demonstrates that plant species exhibit clear trade-offs among their vegetative and reproductive traits and that these trade-offs determine interactions with floral visitors. These trade-offs are differentiated along three axes of trait variation: (i) flower number - flower size, (ii) pollinator dependence and (iii) style length. These reproductive trade-offs helped partly explain the presence of floral visitor interactions, but not their visitation rates. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Moreover, we found that the plant species network roles were best explained by plant size and floral reward related traits.

Despite the low predictive power of the main trait variation axes for broad-level interaction patterns (number of interaction partners and visitation rate), we found changes in the interaction patterns among and within floral visitor guilds across these axes that suggest plant life-history strategies influence plant-pollinator interactions. For example, all floral visitor guilds visited plant species with higher pollinator dependence more frequently, and high pollinator dependence was associated with large floral displays and greater pollen quantities (Fig. 1 and Fig. S6). This trend is consistent with previous studies that show plant species with higher reproductive investment tend to be visited by pollinators more frequently [@hegland2005; @lazaro2013; @kaiser2014]. In regard to the flower number - flower size and style length trade-offs, different pollinator guilds showed contrasting visitation rates across the continuum of trait variation, which could be associated with different pollination syndromes at a macroecological scale. For instance, bees and syrphid flies were clearly associated with opposing life-strategies on PC1 and PC3 (Fig. 3) suggesting possible niche partitioning [@palmer2003; @phillips2020] between these two guilds. However, despite floral rewards not being included in the main analysis because there was insufficient data available, floral reward related traits were among the best at characterising species network roles (Fig. 4). More detailed exploration of reproductive trade-offs in conjunction with floral rewards is needed to help elucidate plant-pollinator associations. In any case, it is worth noting that other local factors such as species relative abundances, surely explain part of the observed variability [@vazquez2007; @encinas2012; @bartomeus2016] that reproductive trade-offs do not.

>>>>

Flowering plants have an astonishing diversity of floral structures [@barrett2002; @schiestl2013] that shape plant-pollinator associations [@fenster2004; @dellinger2020]. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints [REF]. Despite the recent advances in the theoretical understanding of the major floral trade-offs (e.g., floral economic spectrum; REF), we lack empirical evidence that helps to elucidate the spectrum of reproductive trait variation at a global scale as done recently for other vegetative and physiological traits [REF].

Macroecological studies that investigate plant reproductive traits are scarce [@baude2016; @munoz2016; @grossenbacher2017; @moeller2017] and consequently, there is poor understanding of how reproductive traits drive interactions with pollinators at large scales [@sargent2008; @rech2016; @salguero2016; @ruger2018]. Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations [@dehling2016]. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches [@fenster2004; @rosas2014] and trait-matching analyses [@stang2009; @bartomeus2016]. However, plant reproductive traits have been overlooked beyond highly specialised pollination systems [@dellinger2020] and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations [@tur2013; @devaux2014].

@@

Here, we aim to explore the potential trade-offs among reproductive traits and how these influence plant-pollinator interactions. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species’ position in trait-space influence interactions with floral visitors. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species’ roles within networks using a set of complementary interaction network metrics (i.e., number of visits, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks. These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies [@olesen2007; @fortuna2010; @carvalheiro2014] or available in online archives [e.g., The Web of Life, @fortuna2010; and Mangal, @poisot2016]. In total, our network dataset (Table S1) constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative and quantitative form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with the number of visits of floral visitors to individual flowers that account for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency [@ballantyne2015], the number of visits can provide valuable information of the contribution of floral visitors to pollination [@vazquez2005; @vazquez2012]. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version _4.0.3_.

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that _Apis mellifera_ was the floral visitor with the largest proportion of records counted (7.55% of the total). This finding is consistent with previous research showing that _A. mellifera_ was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems [@hung2018]. Hence, to control for the effect of _A. mellifera_ on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding _A. mellifera_. We found that _A. mellifera_, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package _brms_ [@burkner2017]. We modelled presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits ~ PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and number of visits a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with weakly informative priors from the brms function, 3,000 iterations and with previous 1,000 warm up iterations. We set delta ($\Delta$) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function *pp_check* using the _bayesplot_ package [@gabry2019].

Second, to better understand complex trait relationships, we used regression trees. Regression trees are recursive algorithms which can detect complex relationships among predictors and allow identification of the relevance of specific trait combinations on explaining species roles within the network of interaction. We focused exclusively on quantitative traits because almost all categorical traits were statistically associated with the first two axes of trait variation (Table S2). We conducted this analysis using the _rpart_ package [@therneau2015] with method _‘anova’_ with a minimum of 50 observations per terminal node and we used the _rpart.plot_ package [@milborrow2015] to plot the regression trees. We considered the species level indices as response variables (number of visits, normalized degree and specialization) and we performed one regression tree per metric using the different quantitative traits as predictors. We calculated two regression trees per plant species-level metric, one for the full set of species and another for the subset of species for which we had pollen and nectar traits. We focused on regression trees that included floral rewards because they consistently showed pollen and nectar traits as being the best for explaining the different species-level metrics (Fig. S4).

The main axes of trait variation explained partly presence-absence of interaction partners (conditional $R^{2}$ = 0.26; marginal $R^{2}$ = 0.20) but little of the overall number of visits (conditional $R^{2}$ = 0.31; marginal $R^{2}$ = 0.06). However, we found relevant differences across the different floral visitor guilds on both presence-absence of interactions and number of visits (Fig. 3). We found that plants with high flower number and small flowers had higher interaction partners of Coleoptera, non-bee-Hymenoptera and all Diptera guilds but plants with low flower number but large flowers had higher interaction partners of bees and Lepidoptera guilds (flower number - flower size trade-off, PC1; Fig. 3a). However, plant species with high flower number but small flowers had higher number of visits of bees and syrphids guilds (PC1; Fig. 3d). Remarkably, all plant species with higher pollinator dependence had higher number of interacting partners and number of visits for all floral visitor guilds (PC2; Fig. 3b and Fig. 3e). Finally, plant species with short styles and low selfing had higher interaction partners of all guilds but bees that interacted clearly more with plant species with long styles and high selfing (style length trade-off; Fig. 3c). However, for number of visits, we found that plants with long styles and high selfing interacted more frequently with Lepidoptera and non-bee-Hymenoptera guilds (Fig. 3f).

The additional model for both presence-absence of interaction (marginal $R^{2}$ = 0.29; conditional $R^{2}$ = 0.19) and number of visits (marginal $R^{2}$ = 0.30; conditional $R^{2}$ = 0.03) for the most represented families of bees showed that the family Apidae was the main driver of the observed patterns. The contrasting differences between presence-absence of interaction and number of visits for bees on PC1 (Fig. 3a and Fig. 3d) were driven by the family Andrenidae that had higher number of interacting partners but lower number of visits on plant species with low number of large flowers (Fig. S8).

The variance of the different plant species-level network metrics was poorly explained by the three main axes of trait variation (Fig. S9; number of visits ~ PCs, conditional $R^{2}$ = 0.11, marginal $R^{2}$ = 0.02; normalized degree ~ PCs, conditional $R^{2}$ = 0.24, marginal $R^{2}$ = 0.02; and, specialization ~ PCs, conditional $R^{2}$ = 0.37, marginal$R^{2}$ = 0.03). Overall, the most notable trends were found on PC1 and PC3 for number of visits and specialization. On the flower number - flower size trade-off (PC1), number of visits was higher for plant species with more flowers but was lower for plant species with larger flowers (Fig. S9a). On PC1, specialization showed the opposite trend (Fig. S9g). On the style length trade-off (PC3), number of visits was lower for plants with shorter styles and lower autonomous selfing and higher for species with longer styles and higher autonomous selfing (Fig. S9c). Again, specialization showed the opposite trend to number of visits (Fig. S9i).

When we further investigated the combination of traits that drive plant network roles, we found that the regression tree for number of visits was best explained by plant height, nectar concentration and style length (Fig. 4a). Specifically, species taller than 3.9m had the highest number of visits, while species that were shorter than 3.9m and had a nectar concentration lower than 16% had the lowest number of visits. Normalized degree was best explained by nectar concentration, pollen grains per flower, plant height, flower width and autonomous selfing (Fig. 4b). Species with a nectar concentration over 49% had the highest levels of normalized degree, whereas species with nectar concentration lower than 49%, more than 21,000 pollen grains per flower and height less than 0.78m had the lowest normalized degree. Finally, specialization was best explained by plant height, ovule number, pollen grains per flower and autonomous selfing (Fig. 4c). Overall, plant species with the highest specialization were shorter than 1.3m, had more than 14,000 pollen grains per flower and autonomously self-pollinated less than 11% of their fruits. In contrast, species taller or equal than 5.1m and with lower than 14 ovules per flower had the lowest specialization values.

This study demonstrates that plant species exhibit clear trade-offs among their vegetative and reproductive traits and that these trade-offs determine interactions with floral visitors. These trade-offs are differentiated along three axes of trait variation: (i) flower number - flower size, (ii) pollinator dependence and (iii) style length. These reproductive trade-offs helped partly explain the presence of floral visitor interactions, but not their number of visits. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Moreover, we found that the plant species network roles were best explained by plant size and floral reward related traits.

Despite the low predictive power of the main trait variation axes for broad-level interaction patterns (number of interaction partners and number of visits), we found changes in the interaction patterns among and within floral visitor guilds across these axes that suggest plant life-history strategies influence plant-pollinator interactions. For example, all floral visitor guilds visited plant species with higher pollinator dependence more frequently, and high pollinator dependence was associated with large floral displays and greater pollen quantities (Fig. 1 and Fig. S6). This trend is consistent with previous studies that show plant species with higher reproductive investment tend to be visited by pollinators more frequently [@hegland2005; @lazaro2013; @kaiser2014]. In regard to the flower number - flower size and style length trade-offs, different pollinator guilds showed contrasting visitation patterns across the continuum of trait variation, which could be associated with different pollination syndromes at a macroecological scale. For instance, bees and syrphid flies were clearly associated with opposing life-strategies on PC1 and PC3 (Fig. 3) suggesting possible niche partitioning [@palmer2003; @phillips2020] between these two guilds. However, despite floral rewards not being included in the main analysis because there was insufficient data available, floral reward related traits were among the best at characterising species network roles (Fig. 4). More detailed exploration of reproductive trade-offs in conjunction with floral rewards is needed to help elucidate plant-pollinator associations. In any case, it is worth noting that other local factors such as species relative abundances, surely explain part of the observed variability [@vazquez2007; @encinas2012; @bartomeus2016] that reproductive trade-offs do not.

@@

@@

\vspace{5mm}
\blandscape

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=14, fig.width=27,fig.cap= "**Figure 3** Fitted posterior estimates of the presence-absence of interaction (a, b and c) and number of visits (d, e and f) of the different floral visitor guilds in relation to the main axes of trait variation (PC1, PC2 and PC3). PC1 represents the flower number - flower size trade-off, PC2 represents the pollinator dependence trade-off and PC3, the style length trade-off. For visualization purposes, due to large differences between number of visits of bees and the rest of guilds, the number of visits was log-transformed (Y-axis of lower panel)."}

@@

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=4.25, fig.width=12,fig.cap= "**Figure 4** Contribution of traits in plant's network roles. Regression tree analysis of number of visits (log-transformed), normalized degree and specialization for the subset of species with quantitative data for pollen and nectar traits. The superior value inside the node indicates the mean value of the different species-level metric and the lower value, the percentage of species that are considered in each node. Thus, the top node has the mean value of the named trait for the 100% of species. Each node has a yes/no question and when the condition is fulfilled, the branch turns to the ‘yes’ direction and when not, to the ‘no’ direction. This rationale is followed in all the regression trees as indicated in the first branch division of the topmost node of each tree."}

@@

rpart.plot(tree1, box.palette="GnOr", main="(a) Number of visits", cex.main=1.75,tweak = 1.25,type = 2) 
rpart.plot(tree2, box.palette="GnOr", main="(b) Normalized degree", cex.main=1.75,tweak = 1.25,type = 2)
rpart.plot(tree3, box.palette="GnOr", main="(c) Specialization", cex.main=1.75,tweak = 1.25,type = 2)

@@
